Software Quality Assurance in INDIGO-DataCloud project: a converging evolution of software engineering practices to support European Research e-Infrastructures

### 2019-09-17 paper-jgc.tex
remove background section

The paper is organized as follows. %Section~\ref{sec:b} summarizes the essential context of the research problem.
Section~\ref{sec:ev} details the evolution of research on software reliability in the selected EC--projects. 
Section~\ref{sec:ntsr} provides information about new research trends in software reliability based upon the
experiences from the recent {\sl INDIGO--DataCloud} project (INDIGO--DC). Conclusions are drawn in Section~\ref{sec:con}.

@@

One of the fundamental pillars of the {\sl INDIGO--DataCloud}'s SQA process has been established
by the outcomes resulted from similar EC--funded e--Infrastructure--related projects over
the last 15 years. In particular, in this section we are primarily interested in those projects that consisted in
the collaborative development of software solutions, intended for improving the accessibility and exploitation 
of distributed computing technologies by research communities within the European research
landscape. We will highlight the most representative software engineering
practices adopted within each of these projects to understand how they inspired the SQA process 
definition and implementation of {\sl INDIGO--DataCloud}.

The EC projects considered in this paper are listed chronologically in Table~\ref{tab:eup}.
Each project paved the way for the following ones, according to a logic evolution in which
the quality and reliability of the software solutions have progressively increased. As shown in Table~\ref{tab:feat}, the most recent projects are more
committed to, and aware of, software reliability practices.
Table~\ref{tab:feat} also shows the topics addressed in the following sections, highlighting
the commonalities and specific features on software reliability.

@@

Project monitoring and reporting mechanisms allowed to assess the risks, efforts and documentation
produced. Development guidelines \cite{datagridguide} were produced, covering the various phases, including
1) packaging, 2) test and validation, and 3) style and naming convention. The specification of quality indicators
\cite{datagridindicators} facilitated the detection of deviations in the collected metric data. Figure
\ref{fig:efficiency} shows the efficiency performance indicator, measured in production environments as the
number of jobs successfully completed divided by the total number of jobs submitted. As it can be observed,
there is a positive trend of efficiency over the months, with the unique exception of the LCG1 experiment.
\begin{figure}[h]
\centering
%\includegraphics[width=0.8\textwidth, height=50mm]{images/datagridEff.png}
\includegraphics[width=1\textwidth]{images/datagridEff.png}
\caption{Efficiency \% = Number of jobs successfully completed / Total Number of jobs submitted: results are provided for five experiments involved in the project.}
\label{fig:efficiency}
\end{figure}

@@

\subsection{INtegrating Distributed data Infrastructures for Global explOitation}

The {\sl INDIGO--DataCloud} (INtegrating Distributed data Infrastructures for Global
explOitation, Apr 2015 -- Sep 2017) \cite{salomoni2018indigo} is the last
project of software development considered in this paper. The project addressed
the existing gaps at the Cloud Platform--as--a--Service (PaaS) and Software--as--a--Service (SaaS) levels,
helping developers, e--Infrastructure administrators and scientific communities to exploit
Cloud computing benefits.

\section{New trends in software reliability}
\label{sec:ntsr}

{\sl INDIGO--DataCloud} is the most recent project included in this paper, it
leverages the lessons learned from the previous experiences described in
Section~\ref{sec:ev} and, consequently, it will be thoroughly analyzed. The expertise gathered throughout these years was highly
profitable regarding evaluation and application of innovative software engineering
methodologies, management of new technologies to put those methodologies
into practice, compilation and analysis of an appropriate set of metrics to measure the
quality of the implemented solutions.
An additional level of complexity, comes from the fact that these large collaborative projects
have an extensive list of partners.

In the following sections, we describe the three main pillars that sustain the project's strategy to attain reliability throughout the software lifecycle:

@@

@@

>>>>

The paper is organized as follows. %Section~\ref{sec:b} summarizes the essential context of the research problem.
Section~\ref{sec:ev} details the evolution of research on software reliability in the selected EC--projects. 
Section~\ref{sec:ntsr} provides information about new research trends in software reliability based upon the
experiences from the recent {\sl INDIGO--DataCloud} project. Conclusions are drawn in Section~\ref{sec:con}.

@@

One of the fundamental pillars of the {\sl INDIGO--DataCloud}'s SQA process has been established
by the outcomes resulted from similar EC--funded e--Infrastructure--related projects over
the last 15 years. In particular, in this section we are primarily interested in hightlighting the 
collaborative
software development efforts within the category of projects that improved the accessibility and
exploitation of distributed computing technologies in the European research landscape. As such, 
we will present the most representative software engineering practices adopted by those projects
to understand how they ultimately inspired the SQA process definition and implementation in 
{\sl INDIGO--DataCloud}, further discussed in Section~\ref{sec:ntsr}.

The EC projects considered in this paper are listed chronologically in Table~\ref{tab:eup}.
Each project paved the way for the following ones, according to a logic evolution in which
an increasing awareness in software quality and reliability has been showcased. As shown in 
Table~\ref{tab:feat}, the most recent projects are more committed to, and aware of, software
engineering practices. Here we include {\sl INDIGO--DataCloud} (INDIGO--DC) -- see 
Section~\ref{sec:ntsr} -- as the reference project where all the listed features have been achieved.

@@

Project monitoring and reporting mechanisms allowed to assess the risks, efforts and documentation
produced. Development guidelines \cite{datagridguide} were produced, covering the various phases, including
1) packaging, 2) test and validation, and 3) style and naming convention. The specification of quality indicators
\cite{datagridindicators} facilitated the detection of deviations in the collected metric data. Figure
\ref{fig:efficiency} shows the efficiency performance indicator, measured in production environments as the
number of jobs successfully completed divided by the total number of jobs submitted. As it can be observed,
there is a positive trend of efficiency over the months, with the unique exception of the LCG--1 prototype.
\begin{figure}[h]
\centering
%\includegraphics[width=0.8\textwidth, height=50mm]{images/datagridEff.png}
\includegraphics[width=1\textwidth]{images/datagridEff.png}
\caption{Efficiency \% = Number of jobs successfully completed / Total Number of jobs submitted: results are provided for five experiments involved in the project.}
\label{fig:efficiency}
\end{figure}

@@

%\subsection{INtegrating Distributed data Infrastructures for Global explOitation}
%
%The {\sl INDIGO--DataCloud} (INtegrating Distributed data Infrastructures for Global
%explOitation, Apr 2015 -- Sep 2017) \cite{salomoni2018indigo} is the last
%project of software development considered in this paper. The project addressed
%the existing gaps at the Cloud Platform--as--a--Service (PaaS) and Software--as--a--Service (SaaS) levels,
%helping developers, e--Infrastructure administrators and scientific communities to exploit
%Cloud computing benefits.

\section{INDIGO--DataCloud Software Quality Assurance Process}
\label{sec:ntsr}

{\sl INDIGO--DataCloud} project \cite{salomoni2018indigo} %is the most recent project included in this paper, it
leverages the lessons learned from the previous experiences described in
Section~\ref{sec:ev}. The expertise gathered throughout these years was highly
profitable regarding the adoption of state--of--the--art software engineering
methodologies, the management of new technologies to put those methodologies
into practice and the analysis of an appropriate set of metrics to measure the
quality of the implemented solutions. All of which with the additional complexity of
being framed under collaborative environments with an extensive list of partners.

However, new challenges appear in the course of a software development project, which in some
cases are tackled using novel software engineering approaches. In this regard, 
{\sl INDIGO--DataCloud} faced analogous challenges as the ones reported by parallel EC--funded 
software development initiatives \cite{casale2016current} that were addressed using similar 
solutions, such as relying on DevOps practices for tackling the software lifecycle management. 
The SQA process followed by {\sl INDIGO--DataCloud}
combines these solutions with pragmatic approaches built upon external and first--hand 
experiences. In the following sections, we describe the three main pillars that sustain the
project's strategy for attaining quality in the software produced:

@@

As we will describe hereafter, the {\sl INDIGO--DataCloud} SQA process is not restricted to software targeted to distributed
computing environments, but instead oriented to the management of any form of collaboration driven software development, composed
by multiple software projects that collaborate jointly. However, there are particularities, such
as the stage--rollout phase, that are attached to an e--Infrastructure's operation, thus might not
be applicable to all cases.

@@

\bibitem{casale2016current}
G. Casale \emph{et al.},
``Current and future challenges of software engineering for services and applications,''
\emph{Procedia Computer Science},
vol. 97, no. 3, pp. 34--42, 2016.
